%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Do not alter this block (unless you're familiar with LaTeX
\documentclass{article}
\usepackage[margin=1in]{geometry} 
\usepackage{amsmath,amsthm,amssymb,amsfonts, fancyhdr, color, comment, graphicx, environ}
\usepackage{xcolor}
\usepackage{mdframed}
\usepackage[shortlabels]{enumitem}
\usepackage{indentfirst}
\usepackage{hyperref}
\usepackage{algorithm2e}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=blue,
}


\pagestyle{fancy}


\newenvironment{problem}[2][Problem]
    { \begin{mdframed}[backgroundcolor=gray!20] \textbf{#1 #2} \\}
    {  \end{mdframed}}

% Define solution environment
\newenvironment{solution}
    {\textit{Solution:}}
    {}

\renewcommand{\qed}{\quad\qedsymbol}

% prevent line break in inline mode
\binoppenalty=\maxdimen
\relpenalty=\maxdimen

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Fill in the appropriate information below
\lhead{Pengju Zhang}
\rhead{CSC-421} 
\chead{\textbf{Assignment 1}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%problem 1
\begin{problem}{1}
\textbf{[10 pts]}
Given a collection of $n$ nuts and a collection of $n$ bolts, arranged in an increasing order of size, give an $O(n)$ time algorithm to check if there is a nut and a bolt that have the same size. The sizes of the nuts and bolts are stored in the sorted arrays \textit{NUTS[1..n]} and \textit{BOLTS[1..n]}, respectively. Your algorithm can stop as soon as it finds a single match (i.e, you do not need to report all matches).
\end{problem}
\begin{solution}
Since we have two sorted array, we want to keep track both of them with two iterators, \textit{i} for \textit{NUTS[1..n]} and \textit{j} for \textit{BOLTS[1..n]}.
\begin{algorithm}
\SetKwComment{Comment}{/* }{ */}
	$i, j\gets 1$\;
	\While{i $<$ n and j $<$ n} {
		\uIf{NUTS[i] = BOLTS[j]} {return true\Comment*[l]{Once we find two identical number in both array we stop looping}} 
		\uElseIf{NUTS[i] $>$ BOLTS[j]} {$j\gets j+1$\Comment*[l]{We want smaller number array increase their index and compare again with greater one}}
		\Else{$i\gets i+1$\;}
	}
\end{algorithm} 
\newline
We go through index in both array only once, so the worse case is $O(2n) \approx O(n)$. \newline
Noticed that if $\textit{NUTS[1..n]} \cap \textit{BOLTS[1..n]} = \emptyset$, then their range is not overlap and there is no reason to keep searching, thus we can add initial statement.
\begin{algorithm}
	$i, j\gets 1$\;
	\While{i \textless n and j \textless n} {
		\If{NUTS[n] $<$ BOLTS[1] or NUTS[1] $>$ BOLTS[n]} {return;}
		\uIf{NUTS[i] == BOLTS[j]} {return true;}
		\uElseIf{NUTS[i] $>$ BOLTS[j]} {$j\gets j+1$\;}
		\Else{$i\gets i+1$\;}
	}
\end{algorithm}
\end{solution}

%problem 2
\newpage
\begin{problem}{2}
\textbf{[15 pts]} Let $A[1..n]$ be an array of distinct positive integers, and let $t$ be a positive integer.
	\begin{enumerate}
		\item \textbf{[5 pts]} Assuming that $A$ is sorted, show that in $O(n)$ time it can be decided if $A$ 			contains two distinct elements $x$ and $y$ such that $x + y = t$.
		\item \textbf{[10 pts]} Use part (a) to show that the following problem, referred to as the 3-Sum 				problem, can be solved in $O(n^{2})$ time:
		\newline
		(3-SUM) Given an array $A[1..n]$ of distinct positive integers that is not (necessarily) sorted, and a positive 		integer $t$, determine whether or not there are three distinct elements $x, y, z$ in A such that $x + y + z 		= t$.
	\end{enumerate}
\end{problem}
\begin{solution}
\newline
\begin{enumerate}
	\item One solution is to set two $x$ and $y$ be the minimum (A[1]) and maximum (A[n]) of array $A[1..n]$, and then assume that the value of t is sum of $x$ and $y$, if the sum is greater than t we decrease the index of the $y$, and similarly if the sum is less than t we increase the index of the $x$.
	\begin{algorithm}
		$i\gets 1$\, $j\gets n$\;
		\While{i $<$ j} {
			\uIf{A[i] + A[j] = t}{return true;}
			\uElseIf{A[i] + A[j] $>$ t}{$j\gets j-1$\;}
			\Else{$i\gets i+1$\;}
		}
	\end{algorithm}
	\newline
	The worst case is value of $i$ reach $j$ which takes $n-1$ times, thus time complexity is $O(n)$.
	\item By adding extra variable $z$, meaning we want to go through part (a) process for all elements in $A[1..n]$ such that $x + y + z = t$ and $x\neq y \neq z$.
	\begin{algorithm}
		\For{$k\gets 1$ \KwTo $n$}{
			$i\gets 1$\, $j\gets n$\;
			\While{i $<$ j and i $\neq$ k and j $\neq$ k} {
				\uIf{A[i] + A[j] + A[k] = t}{return true;}
				\uElseIf{A[i] + A[j] + A[k]$>$ t}{$j\gets j-1$\;}
				\Else{$i\gets i+1$\;}
			}
		}
	\end{algorithm}
	\newline
	By adding a for loop for part (a), our time complexity is $O(n^2)$ in this case.
\end{enumerate}
\end{solution}

%problem 3
\newpage
\begin{problem}{3}
\textbf{[10 pts]} Let $A[1..n]$ be an array of positive integers ($A$ is not sorted). Pinocchio claims that there exists an $O(n)$-time algorithm that decides if there are two integers in $A$ whose sum is 1000. Is Pinocchio right, or will his nose grow? If you say Pinocchio is right, explain how it can be done in $O(n)$ time; otherwise, argue why it is impossible.
\end{problem}
\begin{solution}
	Yes, by using hash table to search specific value, it only takes $O(n)$ time complexity.
	Assume $A[1..n]$ store in hash table, then we could implement a two sum algorithm as following.
	\begin{algorithm}
		\For{$i\gets 1$ \KwTo $n$}{
			$x\gets 1000 - A[i]$\;
			\If {x exist in A[1..i]} {return true;}
		}
	\end{algorithm}
	\newline
	As mentioned before, we can use hash table search method (A.contains(x) in some languages) to find whether $x$ exist in array, thus it takes $O(n)$ time complexity.
\end{solution}

%problem 4
\newpage
\begin{problem}{4}
\textbf{[10 pts]} Let $A[1..n]$ be an array of points in the plane, where A[i] contains the coordinates $(x_i, y_i)$ of a point $p_i$, for $i = 1,...,n$. Give an $O(n lg n)$ time algorithm that determines whether any two points in $A$ are identical (that is, have the same $x$ and $y$ coordinates).
\end{problem}
\begin{solution}
	Assume by using heapsort (time complexity of $O(nlogn)$) to sort $A[1..2]$ as $A_{sorted}[1..n]$. Then we compare two adjacent points to find if there are two identical coordinates.
	\begin{algorithm}
		$i\gets 1$\;
		\While{i $<$ n - 1} {
			\uIf{$A_{sorted}[i].x = A_{sorted}[i + 1].x$ and $A_{sorted}[i].y = A_{sorted}[i + 1].y$} {
				return true;
			}
			\Else{$i\gets i+1$\;}
		}
	\end{algorithm}	
	\newline
	The total time complexity of cost is $O(nlogn) + O(n) \approx O(nlogn)$.
\end{solution}

%problem 5
\newpage
\begin{problem}{5}
\textbf{[15 pts]} Show how to determine in $O(n^2 lg n)$ time whether any three points in a set of $n$ points are collinear.
\end{problem}
\begin{solution}
	Assume the initial value in $A[1..n]$ and we first want to compute slope of all possibilities.
	\begin{algorithm}
	\SetKwComment{Comment}{/* }{ */}
		$A_{slope}$\Comment*[l]{Initialize an empty array $A_{slope}$ for storing slopes}
		$i\gets 1$\;
		\While{i $<$ n} {
			$j\gets i + 1$\;
			\While{j $<$ n} {
				$slope\gets (A[i].y-A[j].y)/(A[i].x-A[j].x)$\;
				$A_{slope}.add(slope)$\;
				$j\gets j+1$\;
			}
		$i\gets i+1$\;
		}
	\end{algorithm}
	\newline
	Next step by using nested loop for checking two identical slope in $A_{slope}$, iff this happens, three points of set are collinear. Totally the time complexity is $O(n^2)+O(n^2)\approx O(n^2) < O(n^2 lg n)$.
	
	
\end{solution}


\end{document}